\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

\title{CSC165 Assignment 1}
\author{Axel Steingrimsson\\
  999707143,\\
  \texttt{axel.steingrimsson@mail.utoronto.ca}}
\date{\today}						% Activate to display a given date or no date

\begin{document}
\maketitle




\section{Question 1}

\subsection{a)}
C: Carol, D: Dan

$B(C) \land B(D) \Rightarrow P(C, D) \land P(D, C)$

Negation: If Carol and Dan don't play this game, then they haven't played it with each other.

$\neg(B(C) \land B(D)) \Rightarrow \neg(P(C, D) \land P(D, C))$
\subsection{b)}
C: Class, B: Bob

$\forall_x \in C, \neg P(x, B)$

Negation: Everyone in the class has played this 2 person game with Bob

$\forall_x \in C, P(x, B)$

\subsection{c)}
C: Class 

$\exists_x \in C, \neg B(x)$

Negation: There is no one in the class who does not play this 2 person game

$\exists_x \in C, \neg B(x), x = 0$

\subsection{d)}
C: Class

$\exists_x \in C, \exists_y \in C, \exists_z \in C, P(x, y) \Rightarrow P(x, z), y \neq z$

Negation: Everyone in your class who plays this 2-person board game has played the game with only one person. 

$\exists_x \in C, \exists_y \in C, \exists_z \in C, \neg P(x, y) \Rightarrow \neg P(x, z), y = z$


\subsection{e)}
C: Class

$\exists_x \in C, \forall_y \in C, P(x, y)$

Negation:  There is no student in your class has played this 2-person board game with everyone in your class. 

$\exists_x \in C, \forall_y \in C, \neg P(x, y)$

\section{Question 2}
\subsection{a)}
Bob has not traveled outside of the country.
\subsection{b)}
Some of the students have traveled to Denmark, but all of the students have traveled to France.
\subsection{c)}
Pete and Mona have both traveled to some countries, but they have not traveled to the same ones. 
\subsection{d)} %come back to
None of the students have traveled to the same countries. 
\subsection{e)}
There is more than one student that has traveled to all of the countries.
\subsection{f)}
All of the students have been to some countries. 
\section{Question 3}
\subsection{a)}
Converse: If Socrates is Mortal, and all men are Mortal, then Socrates is a man
	
\noindent Contrapositive: If Socrates isn't mortal, and all men are Mortal, then Socrates isn't a man. 
\subsection{b)}
Converse: If the sailing race is held and the lifesaving demonstration
will goes on, then it does not rain or it is not foggy.

\noindent Contrapositive: If the sailing race is not held and the lifesaving demonstration
does not go on, then it rains or it is foggy.
\subsection{c)}
Converse: If I'm sore today, then I was playing hockey yesterday

\noindent Contrapositive: If I'm not sore today, then I didn't play hockey yesterday.
\section{Question 4}
\subsection{a)}
$S1 \Rightarrow S2 $ Because There are no Natural numbers that satisfy S2, but the implication is still valid if S1 is False and S2 is True.
\subsection{b)}
$S3 \Rightarrow S2 $ For the elements in S3, 2 is the only prime number that divides into them, therefore if S3 is true, then S2 is true. 
\subsection{c)}
P: Prime number

$\forall_x \in P, x^n = y \Leftrightarrow y \div x = n, S1 \Rightarrow S4 $ 

There are cases for which S1 can be false, so the statement can remain true if S1 implies S4
\subsection{d)}
$S1 \Rightarrow S5 $

There are an infinite possibility of number which i and j could represent, and S1 does not satisfy that set. 
\section{Question 5}
\subsection{a)}
$((P \Rightarrow Q_1) \land (P \Rightarrow Q_2) \Leftrightarrow (P \Rightarrow (Q_1 \land Q_2))$

$\Leftrightarrow ((P \Rightarrow Q_1) \land (P \Rightarrow Q_2))$

$\Leftrightarrow ((\neg P \lor Q_1) \land (\neg P \lor Q_2))$ Implication 

$\Leftrightarrow (\neg P \lor (Q_1 \land Q_2))$ Distribution Law

$\Leftrightarrow (P \Rightarrow (Q_1 \land Q_2))$ Implication
\subsection{b)}
$ ((\exists_x \in D, P(x) \lor Q(x)) \Rightarrow R(y)) \Leftrightarrow (((\exists_x \in D, P(x) \Rightarrow R(y)) \land ((\exists_x \in D, Q(x)) \Rightarrow R(y)))$

$\Leftrightarrow ((\exists_x \in D, P(x) \lor Q(x)) \Rightarrow R(y))$

$\Leftrightarrow ((\forall_x \in D, \neg P(x) \land \neg Q(x)) \lor R(y))$ Implication

$\Leftrightarrow ((\forall_x \in D, \neg P(x)) \lor R(y)) \land ((\forall_x \in D, \neg Q(x)) \lor R(y))$ Distribution Law

$\Leftrightarrow ((\neg (\exists_x \in D, P(x)) \lor R(y)) \land ((\neg (\exists_x \in D, Q(x)) \lor R(y)))$ DeMorgans Law

$\Leftrightarrow (((\exists_x \in D, P(x)) \Rightarrow R(y)) \land ((\exists_x \in D, Q(x)) \Rightarrow R(y)))$ Implication

\end{document}  



